---
title: 'W6: Functions and Loops'
author: "Ruta Slivkaite"
date: "10/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.

```{r}

## loading libraries
library(tidyverse)
library(gapminder)


## exloring the dataset
head(gapminder_unfiltered)


## steps creating defensive function to calculate GDP of a nation:


  # step 1: defining calc_GDP by assigning it to the output of function. Within the parentheses I indicate the argument names which are going to be used for executing the function.
  # Note: setting default arguments for both, year and country, as NULL means that those arguments will take on those values unless the user specifies otherwise.
  # step 2: defining the body of the function within curly braces ({}).
    # step 2.1: checking whether function arguments are not null values (!is.null) using if statements: it checks whether each additional argument, year and coutry, is set to null, and whenever they’re not null it overwrites the dataset stored in gapminder with a subset given by the non-null argument (a specific country and year).
    # step 2.2: calculating gdp and storing it to a digital object called "gdp"
    # step 2.3: binding the newly calculated gdp value witht the rest of the gapminder data. Sending back the result with return function (optional)


# function

calc_GDP <- function(`gapminder_unfiltered`, year=NULL, country=NULL) {
  if(!is.null(year)) {
    `gapminder_unfiltered` <- `gapminder_unfiltered`[`gapminder_unfiltered`$year %in% year, ]
  }
  if (!is.null(country)) {
    `gapminder_unfiltered` <- `gapminder_unfiltered`[`gapminder_unfiltered`$country %in% country,]
  }
  gdp <- `gapminder_unfiltered`$pop * `gapminder_unfiltered`$gdpPercap

  new <- cbind(`gapminder_unfiltered`, gdp=gdp)
  return(new)
}


## creating the years vector and calculating the GDP of Denmark in those years
years <- c("1967", "1977", "1987", "1997", "2007","2017")
calc_GDP(gapminder, year =  years, country="Denmark")





```

## 2 Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}


## converting gapminder from tibble format to data frame
gapminder <- as.data.frame(gapminder)

## setting the threshold to be used in the loop
lowerThreshold <- 50
upperThreshold <- 70

## steps creating the loop:

  # step 1: iterating through all the countries in the gapminder´s dataframe
  # step 2: testing whether the country starts with letter "B" using startsWith() function
  # step 3: if so, then it must do the following for each of the country starting with letter "B":
    # step 3.1: calculate mean life expectancy and save it to a digital object called "mean_lexp")
    # step 3.2: test whether those countries satisfy any of the if statement condition and print out the corresponding message


# start of the loop
for (i in unique(gapminder$country)) {  
  if (startsWith(i, "B")){
    mean_lexp <- mean(gapminder[gapminder$country == i, "lifeExp"])
    
    
    if(mean_lexp < lowerThreshold){
      cat("Average Life Expectancy in", i, "is less than", lowerThreshold, "\n")
    }
    else if(mean_lexp > lowerThreshold && mean_lexp < upperThreshold){
        cat("Average Life Expectancy in", i, "is between", lowerThreshold, "and", upperThreshold, "\n")
    }
    else{
        cat("Average Life Expectancy in", i, "is greater than", upperThreshold, "\n")
    }
    
}
}
# end of the loop

  
 

```


